### 2015-06-12 tex/table.tex

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
\begin{table}[h]
\label{tab:species_list}
\caption{Colonization-extinction model selected and model performance for each species studied}
\begin{tabular}{@{}llllll@{}}
\toprule
Scientific name             & Common name    & Model & $\Delta$DIC & AUC? & Occurrences \\ \midrule
{\it Abies balsamea}        & Balsam fir     & $T^3 + P^3$ &  & & 18961 \\
{\it Acer rubrum}           & Red maple      &  &  & & 36726 \\
{\it Acer saccharum}        & Sugar maple    &  &  & & 18634 \\
{\it Betula alleghaniensis} & Yellow birch   &  &  & & 10371 \\
{\it Betula papyrifera}     & Paper birch    &  &  & & 15531 \\
{\it Fagus grandifolia}     & American beech &  &  & & 9619 \\
{\it Fraxinus americana}    & White ash      &  &  & & 10385 \\
{\it Picea glauca}          & White spruce   &  &  & & 8526 \\
{\it Pinus banksiana}       & Jack pine      &  &  & & 3023 \\
{\it Populus tremuloides}   & Quaking aspen  &  &  & & 12736 \\ 
{\it Quercus alba}          & White oak      &  &  & & 17280 \\
{\it Ulmus americana}       & American elm   &  &  & & 11597 \\ \midrule
\textbf{Species not started} & & & & & \\
{\it Pinus taeda} & & & & & 16351 \\
{\it Liquidambar styraciflua} & & & & & 14778 \\
{\it Quercus rubra} & & & & & 14337 \\
{\it Liriodendron tulipifera} & & & & & 11045 \\
{\it Quercus velutina} & & & & & 9997 \\
{\it Nyssa sylvatica} & & & & & 9209 \\
{\it Quercus nigra} & & & & & 8291 \\
{\it Pinus strobus} & & & & & 8182 \\
{\it Fraxinus pennsylvanica} & & & & & 7662 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[h]
\label{tab:species_list}
\caption{Parameterized equations for colonization and extinction rate}
\begin{tabular}{lll}
\toprule
Species    & Colonization & Extinction  \\ \midrule
{\it Abies balsamea} & & \\ 
\bottomrule
\end{tabular}
\end{table}