### 2015-06-19 tex/manuscript.tex
Commit everything before I start cleaning up for ms production

\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}

% use proper unicode fonts
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\usepackage{amsmath} % for better display of equations

\usepackage{setspace}
\usepackage{natbib}

\usepackage{titling} % controls the way the title information is displayed
\pretitle{\begin{flushleft}\Large}
\posttitle{\end{flushleft}}
\predate{}
\postdate{}
\preauthor{\begin{flushleft}}
\postauthor{\end{flushleft}}
\setlength{\droptitle}{-3em}

\usepackage{authblk} % adds some nice options for displaying the author list
\renewcommand\Authsep{\protect\\}
\renewcommand\Authands{\protect\\}

%% graphics packages
\usepackage{graphicx}
%\usepackage[nomarkers, tablesfirst]{endfloat} % for final
\usepackage{caption}
%\captionsetup{labelsep=none,textformat=empty} % for final
\captionsetup{labelformat=simple} % for drafts
\usepackage{booktabs}
\usepackage[flushleft]{threeparttable}

%% ----------------------------------
%
%     Title and authorship information
%
%% ----------------------------------

\title{Colonization-extinction dynamics determine tree species range limits in northeastern North America}
\date{}
\author[1,2,3,4]{Matthew V. Talluto (mtalluto@gmail.com)}
\author[1]{Isabelle Boulangeat (isabelle.boulangeat@gmail.com)}
\author[1]{Steve Vissault (steve.vissault@uqar.ca)}
\author[2,3]{Wilfried Thuiller (wilfried.thuiller@ujf-grenoble.fr)}
\author[1]{Dominique Gravel (dominique\_gravel@uqar.ca)}
\affil[1]{Département de biologie, Université du Québec à Rimouski, Rimouski, Quebec, Canada}
\affil[2]{Université Grenoble Alpes, Laboratoire d’Ecologie Alpine (LECA), F-38000 Grenoble, France}
\affil[3]{CNRS, Laboratoire d’Ecologie Alpine (LECA), F-38000 Grenoble, France}
\affil[4]{Author for correspondance. Address: Departament de Biologie, chimie, et geographie, 300, Allée des Ursulines, Rimouski, Quebec G5L 3A1, Canada}

%% ----------------------------------
%
%     END PREAMBLE
%
%% ----------------------------------

\begin{document}
%\doublespacing
%% ----------------------------------
%
%     TITLE PAGE
%
%% ----------------------------------

%TC:ignore
\begin{titlingpage}
	\maketitle

	\begin{flushleft}

	\textbf{Short title:} Colonization-extinction dynamics of eastern North American trees

	\textbf{Keywords:}
	\end{flushleft}
\end{titlingpage}
%TC:endignore

%TC:break abstract
\begin{abstract}
\noindent

 \end{abstract}

%TC:break manuscript
%% ----------------------------------
%
%     INTRODUCTION
%
%% ----------------------------------

%\section*{Introduction}

\section*{Methods}

\subsection*{Model description}

\subsection*{Prevalence model}
We used the RF package in R to build models of regional prevalence for each species. 
For predictors, we initially evaluated a set of xx environmental variables. 
From these, we selected xx variables that were relatively uncorrelated to each other (pairwise Pearson correlation < xx) and that had high loadings in a principal components analysis (PCA) of all variables. 
In this way, we reduced the risk of estimation problems due to correlations among the predictors while retaining variables that were representative of the available climatic space.

\subsection*{Model selection}
For each species, we evaluated a set of 145 colonization-extinction models. 
Each model consisted of at most two environmental variables, one associated with temperature and the other precipitation\. 
The model set was constructed from eight possible model forms (Table?) applied to all possible two-variable combinations of the four energy and four moisture variables (i.e., 16 possible variable combinations), along with two additional single-variable forms for each variable and an intercept-only model. 
Because of the large number of models, fitting all models using MCMC was infeasible; therefore, we performed model selection by computing the Akaike Information Criterion (AIC) on maximum likelihood estimates for each model, fit using simulated annealing (package GenSA in R). 
We then selected the model that minimized AIC for final parameterization.

\subsection*{Parameterization}
We used a Markov Chain Monte Carlo (MCMC) sampler employing a Metropolis-Hastings scheme to estimate posterior distributions for all parameters.
This method has a number of advantages over the simulated annealing approach used for preliminary model selection.
First, MCMC results in full joint posterior distributions of the parameters, allowing for a robust understanding of the uncertainty both in model parameters and predictions.
Second, the use of a Bayesian framework allows the inclusion of prior information.
As our parameters were all estimated on the logit scale, we used the weakly informative Cauchy priors suggested by (cite Gelman paper), with a mean of 0 and (variance?) of 10 for intercept parameters or 2.5 for slope parameters.
Finally, having posterior distributions of the predictors allows for the information produced by our model to be incorporated in future meta-modeling efforts that can further improve our characterization of species' ranges and our understanding of model uncertainty (cite GEB paper).

For each species, we obtained xxx posterior samples...
We assessed convergence...

%\section*{Results}

%\section*{Discussion}

\subsection*{Conclusions}
\begin{itemize}
	\item Range limits of [some/many/most/all] [temperate/boreal] tree species in northeastern north America are determined by colonization-extinction dynamics
	\item Moisture and energy (i.e., mean annual temperature and precipitation) are associated with colonization and extinction rates for many species
	\item For some species, the equilibrium state of the model did not match the predicted distribution of the species, suggesting that these species are either out of equilibrium with the environment, or that other processes (e.g., large-scale disturbances) contribute to their range dynamics.
	\item Resilience (represented by the dominant eigenvalue (c - e) near the transition point) varies widely among species, suggesting greatly differential rates of response to environmental change.
	\item Validation? Model Selection?
\end{itemize}

\subsection*{List of figures}
\begin{itemize}
	\item Fig 1: theoretical/introductory figure or diagram
	\item Fig 2: posterior response curves for c and e w.r.t climate
	\item Fig 3: maps for selected species (with remaining in appendix) - 4 panels - c; e; predicted range; SDM
	\item Fig 4: eigenvalue maps
\end{itemize}

\subsection*{Tables}
\input{table}

\end{document}