### 2015-07-03 tex/tab_si1_col_pars.tex
Added ms support files and a bibliography

\documentclass[11pt]{article}

\usepackage[margin=0.5in]{geometry}
\usepackage{pdflscape}

% use proper unicode fonts
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\usepackage{amsmath} % for better display of equations

\usepackage{caption}
%\captionsetup{labelsep=none,textformat=empty} % for final
\captionsetup{labelformat=simple} % for drafts
\usepackage{booktabs}
\usepackage{multirow}
\usepackage[flushleft]{threeparttable}

\begin{document}

%\newgeometry{margin=1cm}
%\begin{landscape}

\newcommand{\ts}{\scriptsize}
\begin{table}[tb]
\small
\begin{threeparttable}
\label{tab:model_selection}
\caption{Parameterized colonization model selected for each species studied}
\begin{tabular}{lccccccc}
\toprule
\multirow{2}{*}{Scientific name} & \multirow{2}{2cm}{\centering Temperature variable\tnote{*}} & 
\multirow{2}{2cm}{\centering Precipitation variable\tnote{\textdagger}} & \multicolumn{5}{c}{Parameters\tnote{\textdaggerdbl}} \\
& & & $c_0$ & $c_1$ & $c_2$ & $c_3$ & $c_4$ \\ 
\midrule

\multirow{2}{*}{\it Abies balsamea} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & -2.53\tnote{\textsection} & -2.46 & 0.07 & -1.26 & 0.03 \\
 &  &  & {\ts (-2.86, -2.18)} & {\ts (-3.08, -1.81)} & {\ts (-0.01, 0.16)} & {\ts (-1.55, -0.98)} & {\ts (-0.03, 0.09)} \\

\multirow{2}{*}{\it Acer rubrum} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx &  & x.xx \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  & {\ts (x.xx, x.xx)} \\

\multirow{2}{*}{\it Acer saccharum} & \multirow{2}{*}{MAT} & \multirow{2}{*}{PS} & -2.10 & 0.47 & 0.09 & -0.34 &  \\
 &  &  & {\ts (-2.22, -1.99)} & {\ts (0.34, 0.59)} & {\ts (0.01, 0.17)} & {\ts (-0.55, -0.12)} &  \\

\multirow{2}{*}{\it Betula alleghaniensis} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx & x.xx & x.xx & x.xx & \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & \\

\multirow{2}{*}{\it Betula papyrifera} & \multirow{2}{*}{MAT} & \multirow{2}{*}{PS} & x.xx & x.xx & x.xx & x.xx & x.xx \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} \\

\multirow{2}{*}{\it Fagus grandifolia} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx & x.xx &  \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  \\

\multirow{2}{*}{\it Fraxinus americana} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx & x.xx & x.xx & x.xx &  \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  \\

\multirow{2}{*}{\it Fraxinus pennsylvanica} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx & x.xx & x.xx & x.xx & x.xx \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} \\

\multirow{2}{*}{\it Liriodendron tulipifera} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx & x.xx &  \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & \\

\multirow{2}{*}{\it Nyssa sylvatica} & \multirow{2}{*}{MTWQ} & \multirow{2}{*}{PS} & x.xx & x.xx & x.xx & x.xx & x.xx \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} \\

\multirow{2}{*}{\it Picea glauca} & \multirow{2}{*}{MDR} & \multirow{2}{*}{PS} & x.xx & x.xx & x.xx & x.xx & x.xx \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} \\

\multirow{2}{*}{\it Picea mariana} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx &  &  \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  &  \\

\multirow{2}{*}{\it Pinus banksiana} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx & x.xx & x.xx &  &  \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  &  \\

\multirow{2}{*}{\it Pinus strobus} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx &  &  \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  &  \\

\multirow{2}{*}{\it Pinus taeda} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx & x.xx & x.xx & x.xx &  \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & \\

\multirow{2}{*}{\it Populus tremuloides} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & -2.15 & -0.75 & -0.11 & -0.86 &  \\
 &  &  & {\ts (-2.44, -1.86)} & {\ts (-1.48, -0.05)} & {\ts (-0.20, -0.03)} & {\ts (-1.32, -0.43)} &  \\

\multirow{2}{*}{\it Quercus alba} & \multirow{2}{*}{MAT} & \multirow{2}{*}{PS} & x.xx & x.xx & x.xx & x.xx &  \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  \\

\multirow{2}{*}{\it Quercus nigra} & \multirow{2}{*}{MDR} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx & &  \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  &  \\

\multirow{2}{*}{\it Quercus rubra} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx & x.xx & x.xx & x.xx & x.xx \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} \\

\multirow{2}{*}{\it Quercus velutina} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx & x.xx & x.xx \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} \\

\multirow{2}{*}{\it Ulmus americana} & \multirow{2}{*}{MAT} & \multirow{2}{*}{PS} & x.xx & x.xx & x.xx &  & \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & & \\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item [*] {\ts Temperature variable abbreviations: MAT: Mean annual temperature;
		MTWQ: Mean temperature wettest quarter; MDR: Mean diurnal range}
\item [\textdagger] {\ts Precipitation variable abbreviations: MAP: Mean annual precipitation; 
MPWQ: Mean precipitation, warmest quarter; PS: Precipitation seasonality}
\item [\textdaggerdbl] {\ts The parameters fit the following equation for $\gamma$, the 
		colonization rate: $log \left( \frac{\gamma}{1-\gamma} \right) = c_0 + c_1T + c_2P + c_3T^2 + c_4P^2$;
		where $T$ and $P$ are the selected temperature and precipitation variables, 
		respectively, and where a blank cell indicates a parameter that was not included 
		in the best-supported model. 
		All parameters were fit to data scaled to 0 mean and unit variance.}
\item [\textsection] {\ts Values are the mean values from the posterior distribution with 5 and 95\% quantiles}
\end{tablenotes}
\end{threeparttable}
\end{table}

%\end{landscape}
%\restoregeometry

\end{document}