### 2015-07-16 tex/tab_si2_ext_pars.tex

\documentclass[11pt]{article}

\usepackage[margin=0.5in]{geometry}
\usepackage{pdflscape}

% use proper unicode fonts
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\usepackage{amsmath} % for better display of equations

\usepackage{caption}
%\captionsetup{labelsep=none,textformat=empty} % for final
\captionsetup{labelformat=simple} % for drafts
\usepackage{booktabs}
\usepackage{multirow}
\usepackage[flushleft]{threeparttable}

\begin{document}

%\newgeometry{margin=1cm}
%\begin{landscape}

\newcommand{\ts}{\scriptsize}
\begin{table}[tb]
\small
\begin{threeparttable}
\label{tab:model_selection}
\caption{Parameterized extinction model selected for each species studied}
\begin{tabular}{lccccccc}
\toprule
\multirow{2}{*}{Scientific name} & \multirow{2}{2cm}{\centering Temperature variable\tnote{*}} & 
\multirow{2}{2cm}{\centering Precipitation variable\tnote{\textdagger}} & \multicolumn{5}{c}{Parameters\tnote{\textdaggerdbl}} \\
& & & $e_0$ & $e_1$ & $e_2$ & $e_3$ & $e_4$ \\ 
\midrule

\multirow{2}{*}{\it Abies balsamea} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & -3.21\tnote{\textsection} & -1.55 & -0.45 & -0.98 & -0.13 \\
&  &  & {\ts (-3.56, -2.88)} & {\ts (-2.18, -0.94)} & {\ts (-0.54, -0.37)} & {\ts (-1.26, -0.70)} & {\ts (-0.19, -0.07)} \\

\multirow{2}{*}{\it Acer rubrum} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & -3.42 & 0.34 & -0.20 & 0.44 & \\
&  &  & {\ts (3.51, -3.33)} & {\ts (0.28, 0.41)} & {\ts (-0.26, -0.14)} & {\ts (0.36, 0.53)} &  \\

\multirow{2}{*}{\it Acer saccharum} & \multirow{2}{*}{MAT} & \multirow{2}{*}{PS} & -3.18 & 0.73 & 0.47 & 0.32 &  \\
&  &  & {\ts (-3.32, -3.04)} & {\ts (0.61, 0.85)} & {\ts (0.39, 0.55)} & {\ts (0.08, 0.56)} &  \\

\multirow{2}{*}{\it Betula alleghaniensis} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx & x.xx & x.xx & & \\
&  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  & \\

\multirow{2}{*}{\it Betula papyrifera} & \multirow{2}{*}{MAT} & \multirow{2}{*}{PS} & -1.65 & 1.06 & 0.19 &  &  \\
&  &  & {\ts (-1.80 -1.49)} & {\ts (0.89  1.22)} & {\ts (0.14  0.23)} &  &  \\

\multirow{2}{*}{\it Fagus grandifolia} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & -3.50 & 0.00 & -0.20 & 0.64 & 0.23 \\
&  &  & {\ts (-3.68 -3.32)} & {\ts (-0.13  0.12)} & {\ts (-0.33 -0.08)} & {\ts (0.46  0.82)} & {\ts (0.13  0.33)} \\

\multirow{2}{*}{\it Fraxinus americana} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & -2.58 & 0.47 & -0.45 & 0.55 & 0.09 \\
&  &  & {\ts (-2.69 -2.47)} & {\ts (0.32  0.63)} & {\ts (-0.57 -0.34)} & {\ts (0.38  0.73)} & {\ts (-0.01  0.17)}  \\

\multirow{2}{*}{\it Fraxinus pennsylvanica} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & -1.68 & -0.69 & 0.16 & 0.43 & -0.16 \\
&  &  & {\ts (-1.86 -1.51)} & {\ts (-0.98 -0.42)} & {\ts (0.02  0.31)} & {\ts (0.26  0.62)} & {\ts (-0.24 -0.08)} \\

\multirow{2}{*}{\it Liriodendron tulipifera} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx & x.xx &  \\
&  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & \\

\multirow{2}{*}{\it Nyssa sylvatica} & \multirow{2}{*}{MTWQ} & \multirow{2}{*}{PS} & x.xx & x.xx & x.xx &  & x.xx \\
&  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  & {\ts (x.xx, x.xx)} \\

\multirow{2}{*}{\it Picea glauca} & \multirow{2}{*}{MDR} & \multirow{2}{*}{PS} & -2.45 & 0.13 & 0.09 &  & 0.10 \\
&  &  & {\ts (-2.58 -2.31)} & {\ts (0.07  0.19)} & {\ts (-0.01  0.18)} &  & {\ts (0.03  0.16)} \\

\multirow{2}{*}{\it Picea mariana} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx & x.xx &  \\
&  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  \\

\multirow{2}{*}{\it Pinus banksiana} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & -3.30 & -1.32 & -0.62 & -1.17 &  \\
&  &  & {\ts (-4.09 -2.58)} & {\ts (-2.72 -0.02)} & {\ts (-0.86 -0.37)} & {\ts (-1.87 -0.50)} &  \\

\multirow{2}{*}{\it Pinus strobus} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & -3.29 & 0.16 & 0.15 & 0.81 & -0.19 \\
&  &  & {\ts (-3.50 -3.08)} & {\ts (-0.15  0.44)} & {\ts (-0.06  0.36)} & {\ts (0.46  1.16)} & {\ts (-0.39  0.00)} \\

\multirow{2}{*}{\it Pinus taeda} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx & x.xx & x.xx & x.xx &  \\
&  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & \\

\multirow{2}{*}{\it Populus tremuloides} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & -1.40 & 1.20 & -0.06 &  &  \\
&  &  & {\ts (-1.56, -1.25)} & {\ts (1.02, 1.36)} & {\ts (-0.13, 0.01)} &  &  \\

\multirow{2}{*}{\it Quercus alba} & \multirow{2}{*}{MAT} & \multirow{2}{*}{PS} & -3.09 & -0.61 & 0.05 & 1.00 & 0.13 \\
&  &  & {\ts (-3.25 -2.94)} & {\ts (-0.87 -0.32)} & {\ts (-0.10  0.21)} & {\ts (0.81  1.17)} & {\ts (0.02  0.25)} \\

\multirow{2}{*}{\it Quercus nigra} & \multirow{2}{*}{MDR} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx & x.xx &  \\
&  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  \\

\multirow{2}{*}{\it Quercus rubra} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx & x.xx & x.xx & x.xx & x.xx \\
&  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} \\

\multirow{2}{*}{\it Quercus velutina} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & -2.16 & -1.06 & -0.08 & 1.52 &  \\
&  &  & {\ts (-2.27 -2.05)} & {\ts (-1.32 -0.79)} & {\ts (-0.18  0.02)} & {\ts (1.28  1.76)} &  \\

\multirow{2}{*}{\it Ulmus americana} & \multirow{2}{*}{MAT} & \multirow{2}{*}{PS} & -1.59 & -0.16 & -0.24 & 0.26 & \\
&  &  & {\ts (-1.67 -1.51)} & {\ts (-0.27 -0.05)} & {\ts (-0.29 -0.19)} & {\ts (0.15  0.37)} & \\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item [*] {\ts Temperature variable abbreviations: MAT: Mean annual temperature;
		MTWQ: Mean temperature wettest quarter; MDR: Mean diurnal range}
\item [\textdagger] {\ts Precipitation variable abbreviations: MAP: Mean annual precipitation; 
MPWQ: Mean precipitation, warmest quarter; PS: Precipitation seasonality}
\item [\textdaggerdbl] {\ts The parameters fit the following equation for $\epsilon$, the 
		extinction rate: $log \left( \frac{\epsilon}{1-\epsilon} \right) = e_0 + e_1T + e_2P + e_3T^2 + e_4P^2$;
		where $T$ and $P$ are the selected temperature and precipitation variables, 
		respectively, and where a blank cell indicates a parameter that was not included 
		in the best-supported model.
		All parameters were fit to data scaled to 0 mean and unit variance.}
\item [\textsection] {\ts Values are the mean values from the posterior distribution with 5 and 95\% quantiles}
\end{tablenotes}
\end{threeparttable}
\end{table}

%\end{landscape}
%\restoregeometry

\end{document}